\chapter{Assignment \#2: Rankine Cycle 2nd Law Analysis}
\label{ch:ass2}

\section{Background}
Russia has recently placed a nuclear powered barge into operation named the \emph{Akademik Lomonosov.}  This vessel is not self-propelled but is equipped with two pressurized water reactors designated as KLT-40S for providing up to 35 MW of electrical power (each).

We will model the power conversion system for the KLT-40S as a simple Rankine Cycle with a Steam Generator (SG), a Turbine Generator (TG), a condenser and a main feed pump (MFP).  Thermal energy is supplied by primary coolant of a pressurized water reactor (PWR); waste heat is rejected to seawater which is modeled as ordinary water.  Relevant design parameters are given in Table \ref{tab:ass2}
\begin{table}
\begin{tabular}{l c}
\toprule
Parameter & Value \\
\midrule
Primary coolant pressure & 12.7 MPa \\
Primary coolant SG inlet temperature & 317$^{\circ}$C \\
Primary coolant SG outlet temperature & 280$^{\circ}$C \\
SG pressure & 3.8 MPa \\
SG steam outlet temperature & 290$^{\circ}$C \\
Steam mass flow rate & 67 kg/s \\
Dead state temperature and pressure & 5$^{\circ}$C, 101 kPa \\
Condenser pressure & 5 kPa\\
Consenser seawater inlet temperature & 5$^{\circ}$C \\
Seawater cooling temperature rise & 10$^{\circ}$C \\
MFP isentropic efficiency & 65\% \\
TG steam outlet quality & 80\% \\
\bottomrule
\end{tabular}
\caption{\emph{Akademic Lomonosov} Nuclear Steam Supply System key parameters.}
\label{tab:ass2}
\end{table}

\newthought{Analyze the }Rankine cycle described and answer the following questions:
\begin{fullwidth}
\begin{enumerate}
\item What is the isentropic efficiency of the TG?
\vspace{1.0cm}
\item What is the net specific work of the Rankine cycle? [kJ/kg]
\vspace{1.0cm}
\item Verify that the net specific heat is the same as the net specific work.
\vspace{1.0cm}
\item What is the mass flow rate of the primary coolant? [kg/s]
\vspace{1.0cm}
\item What is the mass flow rate of the condenser cooling water? [kg/s]
\vspace{1.0cm}
\item What is the rate of exergy input (to the Rankine cycle) from the primary coolant? [kW]
\vspace{1.0cm}
\item What is the rate of exergy rejection to the environment via condenser cooling water? [kW]
\vspace{1.0cm}
\item Compute the rate of exergy destruction in the Rankine cycle components; fill in the table below:
\begin{table}
\begin{tabular}{l | l}
\toprule
\textbf{Component} & \textbf{Rate of Exergy Destruction [kW]} \\
\hline
Steam generator &  \\
\hline
Turbine generator & \\
\hline
Condenser & \\
\hline
Main feed pump & \\
\hline
\hline
Total Exergy Destruction Rate & \\
\bottomrule
\end{tabular}
\end{table}

\vspace{1.0cm} 

\item What is the rate of exergy transfer through work? [kW]

\vspace{1.0cm}

\item Verify the 2nd law of thermodynamic balance:
$$\text{Exergy in - Exergy out } = \dot{W} + \dot{E}_{d}$$


\end{enumerate}
\end{fullwidth}
